---
title: "Psychophysiological Interaction (PPI): Data Manipulation"
author: "Ekarin Eric Pongpipat, M.A."
date: "2019-01-06"
output: 
  html_document:
    highlight: textmate
    theme: lumen
    code_folding: hide
    code_download: true
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

I believe it is helpful to think of PPI as a essentially *a lot* of data manipulation/wrangling followed by multi-level modeling.

This example will focus on how to perform the data manipulation portion of PPI.

This example assumes that the reader:

1. is familiar with standard task fMRI analyses and jargon
2. has a conceptual understanding of psychophysiological interaction (PPI)
3. has pre-processed the fMRI data

## R Packages

First, let's load some packages.

```{r, warning = F}
packages <- c("tidyverse", "eepR", "furrr", "ggpubr")
xfun::pkg_attach(packages, install = T, message = F)

script_dir <- "~/Documents/GitHub/ppi/R/"
script_files <- list.files(script_dir)
script_path <- paste0(script_dir, script_files)
script_ppi <- lapply(script_path, source)
```

## Define Parameters

Let's define some MRI parameters and an upsample factor for convolution and deconvolution.

In this example, we are going to use SPM's first gamma variate (`spmg1`) for the HRF. This is SPM's default HRF other current options include `gam` and `block`.

### MRI
```{r}
tr <- 1.5
n_volumes <- 260
```

### Files
```{r}
events_file <- "~/Box/my_mri/behavioral/nback_3tb4188_2017_Sep_13_1341_soa.csv"
out_file <- NULL
```

### PPI Options
```{r}
hrf_name <- "spmg1"
upsample_factor <- 16
detrend_factor <- 2

psy_contrast_table <- cbind(nback_vs_baseline = c(1, 1, 1, 1, -4)/5,
                                              task_vs_control = c(-3, 1, 1, 1, 0)/4,
                                              linear_task = c(0, -1, 0, 1, 0),
                                              quadratic_task = c(0, -1, 2, -1, 0)/3)
psy_contrast_table
```

## Ideal HRF

Let's first obtain the hemodynamic response function (HRF) that we are going to use during convolution and deconvolution steps. 

```{r, warning = F, message = F}
hrf <- get_hrf_afni(hrf_name, tr, upsample_factor)
```

## Psychological Variables

### Load Data
```{r}
df_psy <- read.csv(events_file) %>%
  rename(onset = start, trial_type = block) %>%
  mutate(duration = ifelse(trial_type == "0-back", 10 * 2.5, 20 * 2.5)) %>%
  select(run, onset, duration, trial_type) %>%
  group_by(run) %>%
  nest() %>%
  mutate(data = map(data, as.data.frame))
```

### Create
```{r}
psy_var <- apply(df_psy[, "data"], 1, function(x) create_psy_var(x[[1]], psy_contrast_table, hrf, tr, n_volumes, upsample_factor))
```

## Physiological Variable

### Load Data
```{r, fig.width = 12, fig.height = 6}
dir_phys <- "/Volumes/shared/KK_KR_JLBS/Wave1/MRI/FMRI/PPI/Nback_individual_covariates/PHYS_MNI_42_-42_42/3tb1780/"
files_phys <- list.files(dir_phys, "_mean.1D")
path_phys <- paste0(dir_phys, files_phys)
df_phys <- tibble(file = files_phys,
                  path = path_phys) %>%
  mutate(run = str_remove(file, "_mean.1D"),
         run = str_remove(run, "r"),
         data = future_map(path, function(x) read.csv(x, header = F, col.names = "data")))
```

### Create
```{r}
phys_var <- apply(df_phys[, "data"], 1, function(x) create_phys_var(x[[1]], detrend_factor, upsample_factor, hrf))
```

## Psychophysiological Interaction Variable

### Create
```{r}
temp_psy_var <- map(psy_var, "upsample")
temp_phys_var <- map(phys_var, "deconvolve")
ppi_var <- list()
for (i in 1:3) {
  ppi_var[[i]] <- create_ppi_var(temp_psy_var[[i]], temp_phys_var[[i]], hrf, tr, n_volumes, upsample_factor)
}
```

## Design Matrix

### Initialize
```{r}
temp_psy_var <- map(psy_var, "downsample")
temp_phys_var <- map(phys_var, "detrend")
temp_ppi_var <- map(ppi_var, "downsample")

df_design_mat <- tibble(run = c(1:3)) %>%
  mutate(data = future_map(run, function(x) create_design_matrix(temp_psy_var[[x]], temp_phys_var[[x]], temp_ppi_var[[x]]))) %>%
  unnest()

head(df_design_mat)
```

```{r}
func_fig_data_heat_map <- function(data) {
  colnames(data) <- paste0(str_pad(1:ncol(data), 2, "left", "0"), "_", colnames(data))
  apply(data, 2, scale_min_max) %>%
    as.data.frame() %>%
    mutate(volume = row_number()) %>%
    gather(., variable, value, -volume) %>%
    ggplot(., aes(variable, volume, fill = value)) +
    geom_raster() + 
    scale_fill_distiller(palette = "Greys", direction = 1) + 
    theme_minimal() +
    theme(axis.text.x = element_text(hjust = 1, angle = 45)) +
    labs(x = NULL)
}

func_fig_data_ts_long <- function(data) {
  colnames(data) <- paste0(str_pad(1:ncol(data), 2, "left", "0"), "_", colnames(data))
  data %>%
  mutate(volume = row_number()) %>%
  gather(., variable, value, -volume) %>%
  ggplot(., aes(volume, value)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 1) +
  theme_minimal()
}

func_fig_data_heat_map(df_design_mat)
```

```{r, fig.width = 8, fig.height = 12}
func_fig_data_ts_long(df_design_mat)
```

### Save
```{r}
if (!is.null(out_file)) {
  write.csv(df_design_mat_final, out_file)
}
```
