---
title: "Psychophysiological Interaction (PPI): Data Manipulation"
author: "Ekarin Eric Pongpipat, M.A."
date: "2019-01-06"
output: 
  html_document:
    highlight: textmate
    theme: lumen
    code_folding: hide
    code_download: true
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

I believe it is helpful to think of PPI as a essentially *a lot* of data manipulation/wrangling followed by multi-level modeling.

This example will focus on how to perform the data manipulation portion of PPI.

This example assumes that the reader:

1. is familiar with standard task fMRI analyses and jargon
2. has a conceptual understanding of psychophysiological interaction (PPI)
3. has pre-processed the fMRI data

## R Packages

First, let's load some packages.

```{r, warning = F}
packages <- c("tidyverse", "eepR", "furrr", "ggpubr")
xfun::pkg_attach(packages, install = T, message = F)

script_dir <- "~/Documents/GitHub/ppi/R/"
script_files <- list.files(script_dir)
script_path <- paste0(script_dir, script_files)
script_ppi <- lapply(script_path, source)
```

## Define Parameters

Let's define some MRI parameters and an upsample factor for convolution and deconvolution.

In this example, we are going to use SPM's first gamma variate (`spmg1`) for the HRF. This is SPM's default HRF other current options include `gam` and `block`.

```{r}
tr <- 1.5
n_volumes <- 260
hrf_name <- "spmg1"
upsample_factor <- 16
detrend_factor <- 2

events_file <- "~/Box/my_mri/behavioral/nback_3tb4188_2017_Sep_13_1341_soa.csv"

psy_contrast_table <- cbind(nback_vs_baseline = c(1, 1, 1, 1, -4)/5,
                                              task_vs_control = c(-3, 1, 1, 1, 0)/4,
                                              linear_task = c(0, -1, 0, 1, 0),
                                              quadratic_task = c(0, -1, 2, -1, 0)/3)
psy_contrast_table

out_file <- NULL
```

## Ideal HRF

Let's first obtain the hemodynamic response function (HRF) that we are going to use during convolution and deconvolution steps. 

```{r, warning = F, message = F}
hrf <- get_hrf_afni(hrf_name, tr, upsample_factor)

ggplot(hrf %>% mutate(volume = row_number()), aes(volume, hrf)) +
  geom_line() +
  theme_minimal() +
  labs(title = paste0("HRF: ", toupper(hrf_name)),
       subtitle = "Upsample: ", upsample_factor)
```

## Psychological Variables

### Load Data
```{r}
df_psy <- read.csv(events_file) %>%
  rename(onset = start, trial_type = block) %>%
  mutate(duration = ifelse(trial_type == "0-back", 10 * 2.5, 20 * 2.5)) %>%
  select(run, onset, duration, trial_type) %>%
  group_by(run) %>%
  nest()
```

### Create
```{r}
psy_var <- apply(df_psy[,2], 1, function(x) create_psy_var(x[[1]], psy_contrast_table, hrf, tr, n_volumes, upsample_factor))
```

## Physiological Variable

### Load Data
```{r, fig.width = 12, fig.height = 6}
func_fig_phys <- function(data) {
  colnames(data) <- "data"
  
  data <- as.data.frame(data)
  
  fig <- data %>%
    mutate(volume = row_number()) %>%
    ggplot(., aes(volume, data)) +
    geom_line() +
    theme_minimal()
  return(fig)
}

dir_phys <- "/Volumes/shared/KK_KR_JLBS/Wave1/MRI/FMRI/PPI/Nback_individual_covariates/PHYS_MNI_42_-42_42/3tb1780/"
files_phys <- list.files(dir_phys, "_mean.1D")
path_phys <- paste0(dir_phys, files_phys)
df_phys <- tibble(file = files_phys,
                  path = path_phys) %>%
  mutate(run = str_remove(file, "_mean.1D"),
         run = str_remove(run, "r"),
         data = future_map(path, function(x) read.csv(x, header = F, col.names = "data")),
         fig = future_map(data, func_fig_phys),
         fig = future_map2(fig, paste0("Run: ", run), func_add_fig_title))

ggarrange(plotlist = df_phys$fig, ncol = 1)
```

### Create
```{r}
phys_var <- create_phys_var(df_phys, detrend_factor, upsample_factor, hrf)
```

## Psychophysiological Interaction Variable

### Load Data
```{r, , fig.width = 12, fig.height = 12}

```

### Create
```{r}
ppi_var <- create_ppi_var(map(psy_var, "upsample"), map(phys_var, "deconvolve"), hrf, tr, n_volumes, upsample_factor)
```

## Design Matrix

### Initialize
```{r}
df_design_mat <- tibble(run = c(1:3),
                        data = NA)

for (i in 1:nrow(df_design_mat)) {
  
  temp_psy <- df_psy_downsample$data_downsample[[i]]
  temp_phys <- df_phys_detrend$data_detrend[[i]] %>% rename(phys = residual)
  temp_ppi <- df_ppi_downsample$data_downsample[[i]]
  colnames(temp_ppi) <- str_replace(colnames(temp_ppi), "psy_", "ppi_")

  df_design_mat$data[i] <- list(cbind(temp_psy, temp_phys, temp_ppi))
}

df_design_mat <- df_design_mat %>%
  unnest()

func_fig_data_heat_map <- function(data) {
  colnames(data) <- paste0(str_pad(1:ncol(data), 2, "left", "0"), "_", colnames(data))
  apply(data, 2, scale_min_max) %>%
    as.data.frame() %>%
    mutate(volume = row_number()) %>%
    gather(., variable, value, -volume) %>%
    ggplot(., aes(variable, volume, fill = value)) +
    geom_raster() + 
    scale_fill_distiller(palette = "Greys", direction = 1) + 
    theme_minimal() +
    theme(axis.text.x = element_text(hjust = 1, angle = 45)) +
    labs(x = NULL)
}

func_fig_data_ts_long <- function(data) {
  colnames(data) <- paste0(str_pad(1:ncol(data), 2, "left", "0"), "_", colnames(data))
  data %>%
  mutate(volume = row_number()) %>%
  gather(., variable, value, -volume) %>%
  ggplot(., aes(volume, value)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 1) +
  theme_minimal()
}

func_fig_data_heat_map(df_design_mat)
```

```{r, fig.width = 8, fig.height = 12}
func_fig_data_ts_long(df_design_mat)
```

### Add nuisance variables

```{r}

```

### Final
```{r}
df_design_mat_final <- df_design_mat %>%
  group_by(run) %>%
  nest() %>%
  mutate(data = future_map(data, function(x) x %>% mutate(time_linear = row_number(),
                                                          time_linear = scale(time_linear),
                                                          time_quadratic = time_linear^2))) %>%
  unnest() %>%
  ungroup()

func_fig_data_heat_map(df_design_mat_final)
```

```{r, fig.width = 8, fig.height = 12}
func_fig_data_ts_long(df_design_mat_final)
```

### Save
```{r}
if (!is.null(out_file)) {
  write.csv(df_design_mat_final, out_file)
}
```
